README.md

### 2024-11-27 README.Rmd
Zenodo link ⛓️
updated README and GitHub manuscript 📖

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.6077618.svg)](https://doi.org/10.5281/zenodo.6077618)

![Antigua fábrica de La Cerámica Gijonesa](ceramica.JPG)

# Anthropogenic plant communities of the Iberian Atlantic ecoregion

This repository stores all the information related to the manuscript [*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

@@

* *Questions* Which is the diversity of anthropogenic plant communities at the ecoregional level in a transitional territory between temperate and Mediterranean Europe? How are these communities composed by native plants, archaeophytes and neophytes? What are the main environmental and disturbance drivers of the diversity in anthropogenic vegetation?

* *Location* Iberian Atlantic ecoregion, Iberian Peninsula, south-western Europe.

* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic alliances. We used modified TWINSPAN and semi-supervised classification to classify the plots into the revised alliances. We determined the proportion of natives, archaeophytes and neophytes. We described the alliances in terms of vegetation height, flowering phenology and ecological requirements for temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.

* *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*). The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes and 7% neophytes. The main ecological gradients behind this diversity went from dry-sunny to moist-shady habitats, followed by a gradient of low to high disturbance frequency and severity.

* *Conclusions* The human-made habitats of the Iberian Atlantic ecoregion host one third of the ecoregional species pool and one fifth of the Iberian flora. Wet anthropogenic habitats are most affected by neophyte invasions. Annual vegetation of arable lands is a refuge for archaeophytes. Mesic perennial ruderal vegetation is specially rich in native species and can be a biodiversity asset in urban and peri-urban settings.

@@

> Fernández-Pascual E, González-García, Ivesdal G, Lázaro-Lobo A, Jiménez-Alfaro B. [*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.

>>>>

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.11613571.svg)](https://zenodo.org/doi/10.5281/zenodo.11613571)

![](ceramica.JPG)

# Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula

This repository stores all the information related to the manuscript [*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

@@

* *Questions* Can we reconcile regional and European classifications of anthropogenic plant communities at the biogeographical scale? How are these communities characterized by species origins, traits and ecological preferences?

* *Location* Atlantic territories in the NW Iberian Peninsula (a.k.a. Cantabrian Mixed Forests ecoregion); south-western Europe.

* *Methods* We classified 2,508 plots with the aim of being consistent with regional phytosociological expertise, while matching that expertise with current EuroVegChecklist alliances. We used modified TWINSPAN to revise the original phytosociological classification, followed by semi-supervised re-classification of the whole dataset. We determined the proportion of natives, archaeophytes and neophytes. We also described the alliances in terms of species traits (lifeforms, height and flowering phenology) and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and severity).

* *Results* We assigned 2,086 vegetation plots to 25 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*). The plots included 1,149 species: 78% natives, 15% archaeophytes and 7% neophytes. Vegetation groups were organized along a principal axis of abiotic stress (dry-sunny to moist-shady habitats) and a secondary axis of disturbance.

* *Conclusions* In the Iberian Atlantic territories, anthropogenic habitats host one third of the regional plant species pool and one fifth of the Iberian flora. Mesic perennial ruderal vegetation is especially rich in native species and can be a biodiversity asset in urban landscapes. Our biogeographical-level synthesis can improve the management of anthropogenic plant communities and contribute towards a European-level synthesis of human-made vegetation.

@@

> Fernández-Pascual E, González-García, Ivesdal G, Lázaro-Lobo A, Jiménez-Alfaro B. [*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.
